Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-11 paper/ms.tex
update ms

\subsection{Typical and extreme CQVs}
\label{subsec:extreme}
Figure~\ref{fig:catalogscatter} shows some of the derived properties
for our CQV catalog, and contrasts them against both the target star
sample (top panels), and the population of rotating stars in the Pleiades
from \citet{2016AJ....152..114R}.
Generally, CQVs are much younger than typical field stars. 
In terms of their rotation, relative to the Pleiades,
they are among the more rapidly rotating half of M dwarfs.
The dearth of $T$$>$15.5 CQVs in our sample, despite our search extending
to $T$$=$16, is a consequence of the 120-second selection function,
not detection sensitivity.

@@

The lowest mass ($\approx 0.12$\,$M_\odot$) belongs to TIC~267953787.
The catalog contains a few other stars with similar mass.  Given the
small number of sub-stellar mass objects in our target sample, this
suggests that future studies of brown dwarf photometric variability
might also yield complex quasiperiodic variables, though there could
be degeneracies in interpretation with planetary surface features such
as clouds and latitudinal bands
\citep[e.g.][]{2021ApJ...906...64A,2022ApJ...924...68V}.

The most massive CQVs in our sample might represent an interesting new
regime.  To date, the only stars reported to show the CQV phenomenon
are M dwarfs, with typical stellar masses $\lesssim$0.3\,$M_\odot$
\citep{2017AJ....153..152S,2022AJ....163..144G}.  TIC~405754448 and TIC~405910546 however
appear to have masses of $\approx$0.82 and $\approx$0.60\,$M_\odot$
respectively.  The next-highest masses are $\approx$0.40\,$M_\odot$.
The masses for the former two objects are consistent with their
locations in the color--absolute magnitude diagram, and with their
membership in LCC.  Based on its light curve morphology, TIC~405910546
should be studied in greater depth, to confirm it is not an eclipsing
binary.  TIC~405754448 similarly shows distinct morphology from many
of the CQVs in Figure~\ref{fig:cqvs}, in that it has some of the
lowest amplitude dips.  While TIC~405754448 might be an unresolved
binary (RUWE$_{\rm DR3}$=4.7),  the depth of the coherent dips present
in the Sector~11 light curve imply that the CQV signal is indeed from
the apparent primary of the system.  Taken in sum, both of these
objects seem to suggest that the CQV phenomenon extends up in mass to
pre-main-sequence K dwarfs.

@@

\subsubsection{The lifetime constraint}
Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from e.g.
\citet{2022AJ....164...80R} and references therein, it seems plausible
that CQV occurrence decreases in time, bottoming to zero well before
the $\approx$700\,Myr age of Praesepe.  This is odd in the context of
the prominence scenario, because
pre-main-sequence M dwarfs spin up over the first $\approx$$10^8$\,yr;
prominences might even be expected to be {\it more} common at the age
of the Pleiades than for younger stars.  This broadly assumes that any
star that can support prominences will, and that the magnetic topology
of rapidly rotating M dwarfs at 100\,Myr is similar to that of 
rapidly rotating M dwarfs at say, 10\,Myr.  Based on the degree to
which M dwarfs remain rapid rotators over the first 700\,Myr
\citep[e.g. Figure~13 of][]{2022AJ....164...80R}, it seems surprising
in the gas clump (prominence) scenario that very few stars older than
150\,Myr have been seen to exhibit the phenomenon.  In the dust clump
scenario this is not a problem, because if the dust were externally
sourced, it would be expected to have a finite supply.

>>>>

\subsection{Typical and extreme CQVs}
\label{subsec:extreme}
Figure~\ref{fig:catalogscatter} shows some of the derived properties for our
CQV catalog, and compares them against both the target star sample (top
panels), and against stars in the Pleiades
\citep{2016AJ....152..114R}.  Typical CQV masses span
0.1-0.4\,$M_\odot$, and typical ages span 2-150\,Myr.  This
is a mass and age regime that includes fully convective stars
($M_\star$$\lesssim$0.25\,$M_\odot$), as well as stars with radiative cores
($M_\star$$\gtrsim$0.25\,$M_\odot$; \citealt{2018A&A...619A.177B}).  We found
no obvious morphology differences between the two classes.  In terms of their
rotation, relative to the Pleiades, the CQVs are among the more rapidly
rotating half of M dwarfs.  The dearth of $T$$>$15.5 CQVs in our sample,
despite our search extending to $T$$=$16, is a consequence of the 120-second
selection function, not detection sensitivity.

@@

The lowest mass ($\approx 0.12$\,$M_\odot$) belongs to TIC~267953787.
The catalog contains a few other stars with similar mass.  Given the
small number of sub-stellar mass objects in our target sample, 
future studies of brown dwarf photometric variability
might also yield CQVs, though there could
be degeneracies in interpretation with planetary surface features such
as clouds and latitudinal bands
\citep[e.g.][]{2021ApJ...906...64A,2022ApJ...924...68V}.

The two most massive CQVs in our sample might represent an interesting
new regime, but both have ambiguities.  To date, the only stars
reported to show the CQV phenomenon are M dwarfs, with typical stellar
masses $\lesssim$0.3\,$M_\odot$
\citep{2017AJ....153..152S,2022AJ....163..144G}.  TIC~405754448 and
TIC~405910546 however appear to have masses of $\approx$0.82 and
$\approx$0.60\,$M_\odot$ respectively.  The next-highest masses in our
sample are $\approx$0.40\,$M_\odot$.  The masses for the former two
objects are consistent with their locations in the color--absolute
magnitude diagram, and with their membership in LCC.  However based on
its light curve morphology, TIC~405910546 should be studied in greater
depth, to confirm it is not an eclipsing binary.  For TIC~405754448,
we expect that the source is an unresolved binary, because RUWE$_{\rm
DR3}$=4.7, and because the raw light curves in Sectors 11, 37, and 38
for this source show an additional $\approx$6.5\,day, $\approx$0.3\%
amplitude sinusoidal signal suggestive of an unresolved photometric
companion.  If TIC~405754448 is a K+M binary, then the flux ratio
between the primary and secondary would be expected to be
$\approx$10:1.  It would be somewhat challenging for the M dwarf to
produce the observed 12.9\,hr signal with such a flux deficit, but not
impossible.  Therefore while both of these objects seem to suggest
that the CQV phenomenon may extend up in mass to pre-main-sequence K
dwarfs, more data are needed to fully substantiate this claim.

@@

\subsubsection{The lifetime constraint}
Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from e.g.
\citet{2022AJ....164...80R} and references therein, it seems plausible
that CQV occurrence decreases in time, bottoming to zero well before
the $\approx$700\,Myr age of Praesepe.  This is odd in the context of
the prominence scenario, because
pre-main-sequence M dwarfs spin up over the first $\approx$$10^8$\,yr;
prominences might even be expected to be {\it more} common at the age
of the Pleiades than for younger stars.  This broadly assumes that any
star that can support prominences will, and that the magnetic topology
of rapidly rotating M dwarfs at 100\,Myr is similar to that of 
rapidly rotating M dwarfs at say, 10\,Myr.  Based on the degree to
which M dwarfs remain rapid rotators over the first few gigayears
of their lives
\citep[e.g.][]{2022AJ....164...80R,2022ApJ...936..109P}, it seems surprising
in the gas clump (prominence) scenario that very few stars older than
150\,Myr have been seen to exhibit the phenomenon.  In the dust clump
scenario this is not a problem, because if the dust were externally
sourced, it would be expected to have a finite supply.
